---
title: "hyptests2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hyptests2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
options(repos=c(CRAN="https://cran.r-project.org"))
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The `hyptests2` package provides a suite of tools for performing statistical tests such as T-tests, Chi-square tests, and ANOVA while generating clear visualizations of the data and test results. This package aims statistical tests more accessible for users who are new statistical testing, providing them with graphical interpretations and desired results.

## Motivation

Traditional statistical tests in R produce numeric outputs, which can be difficult to interpret without prior experience. The `visualStats` package enhances these tests by including:

1. **Dynamic visualizations** of the input data and test results.
2. **Clear and interactive workflows** through the Shiny gadget.
3. **Integration with ggplot2** for customizable and high-quality visual output.

This vignette will walk through the core functionality of `hyptests2` and demonstrate how it can be applied.

---

# Installation

The `hyptests2` package can be directly installed from GitHub using the `devtools` package:

```{r, results='hide'}
# Install devtools if not already installed
install.packages("devtools")

# Install hyptests2 from GitHub
devtools::install_github("sofsteph/hyptests2", build_vignettes = TRUE)

```

---

# Core Functions and Examples

The `hyptests2` package includes three main functions: `t_test_2.0`, `chi_square.2.0`, and `anova.2.0`. These functions perform standard statistical tests and include visualizations to aid interpretation.

## Example 1: T-Test with Visualizations

### Dataset
We'll create a simple dataset to perform a T-Test:

```{r, warning=FALSE}
# Simulated dataset
set.seed(123)
group1 <- rnorm(30, mean = 5, sd = 1)
group2 <- rnorm(30, mean = 6, sd = 1.5)

# load the package
library(hyptests2)

# Perform the T-Test
result <- t_test_2.0(group1, group2)
```

### Output
The `t_test_2.0` function outputs:

1. The standard t-test results.
2. A **boxplot** comparing the two groups.
3. A **p-value visualization** showing the t-distribution and the critical regions.

---

## Example 2: Chi-Square Test with Visualizations

### Dataset
Suppose we have the following contingency table:

```{r, warning=FALSE}
observed <- matrix(c(30, 10, 20, 40), nrow = 2, byrow = TRUE,
                   dimnames = list(c("Group 1", "Group 2"),
                                   c("Outcome A", "Outcome B")))

# Perform the Chi-Square Test
result <- chi_square.2.0(observed)
```

### Output
The `chi_square.2.0` function outputs:

1. The standard chi-square test results.
2. A **heatmap** of the observed frequencies.
3. A **p-value visualization** of the chi-square distribution.

---

## Example 3: ANOVA with Visualizations

### Dataset
For ANOVA, let's consider the `iris` dataset:

```{r, warning=FALSE}
# Perform ANOVA
result <- anova.2.0(data = iris, formula = Sepal.Length ~ Species)
```

### Output
The `anova.2.0` function outputs:

1. The ANOVA table.
2. A **boxplot** of the groups.
3. A **p-value visualization** showing the F-distribution.

---

# Using the Shiny Gadget

The `hypGadget()` function provides the user with an interactive interface for performing these tests, and can be launched with the following:

```{r, eval=FALSE}
# Launch the gadget
hypGadget()
```

1. Upload or select your dataset.
2. Choose the statistical test you want to perform.
3. Specify the parameters (e.g., groups for a t-test).
4. View the results and visualizations.

---

# Future Work

1. **Enhanced Interactivity:**
   - Allow users to dynamically filter data in the Shiny gadget.
   - Provide options for customizing ggplot2 themes.

2. **Support for Additional Tests:**
   - Add support for non-parametric tests such as the Wilcoxon and Kruskal-Wallis tests.
   - Add visualization for p-values. 
   - Include correlation and regression analyses with visualizations.

3. **Advanced Visualization Tools:**
   - Include interactive graphics with tools from the `shiny` package.

---

# Conclusion

The `hyptests2` package enhances statistical analysis in R by combining robust methods with intuitive visualizations. This package can help make statistical results more accessible and actionable for students, data analysts, and others.

